Analysis and Figures for Gallego et al, 2020

### 2020-09-24 Manuscript/What.lies.ahead.reviewed.tex
Rmarkdown shows consistency and figures are good to go. Latex document updated

Despite geographic proximity and similar overall species composition (127 of the 222 planktonic taxa were found in both regions and accounted for 98\% of the sequences), the areas under study -- San Juan Island and the Hood Canal -- varied widely in pH, temperature, and other environmental parameters (Figure 1C), with a smooth gradient in conditions along the Hood Canal, and San Juan Island more closely resembling full marine conditions. Different points along the environmental continuum simultaneously showed differences equivalent to those predicted between present-day and future oceans \cite{bopp2013multiple}.

Metabarcoding analysis of eDNA samples generated more than 52.5M sequences from 745 samples. These samples represented biological and technical replicates from 86 unique sampling events. After bioinformatics quality-control (See Methods) the dataset included $\sim${}41M sequences, from 4782 unique amplicon sequence variants (ASVs). Of these, 1374 ASVs (22.8M reads) could be annotated to a taxonomic level of Family or lower. These $\sim$ 500 taxa from 43 phyla were split according to their natural history and habitat (benthic vs. planktonic; see Supporting Information). Because we expect planktonic taxa to vary with water mass \cite{kelly2018tides} and therefore with bottle-sampled carbonate chemistry, here we focus on only the planktonic taxa (N = 222). These taxa showed a seasonal richness gradient between study areas, consistent with documented biodiversity clines in the area (Fig. 1B.1) \cite{dethier_decadal_2008}.

Bray-Curtis dissimilarities among samples revealed large differences in metabarcoding communities due to geographic Area (Hood Canal vs. San Juan Island; F-model = 5.4; p $<$ 0.0001). We therefore performed a constrained analysis of principal components (CAP) for samples within each Area, showing the differences among communities as a function of temperature, pH, and salinity (Fig. 2; for clarity, results for Hood Canal shown; full analysis in the Supporting Information).

@@

These communities were distinguished by changes in the relative abundances of a wide variety of taxa. In the Hood Canal for example, the colder water and higher-pH community (community 1 in Fig. 2C) showed higher eDNA indexes of diatoms \textit{Thalassiosira nordenskioeldii}, \textit{Nitzschia cf. promare}, and \textit{Ditylum brightwellii} -- among other species -- relative to a higher-temperature community (community 3 in Fig. 2C). Communities 2 and 3 had relatively higher values for taxa from a mix of groups including \textit{Emiliania huxleyi}, \textit{Minutocellus polymorphus}, \textit{Micromonas pusilla}, and \textit{Bathycoccus prasinos}. Community 3 (Fig. 2A,B) and similar planktonic communities occupy the spectrum of environmental conditions most likely to be encountered in near-future climate scenarios as temperature rises and pH falls. For example, we expect the Hood Canal in 2095 to have the conditions in which community 3 is the most likely community 75\% of the time, an increase of 22\% compared to 2017(see Supplemental Material).

@@

\begin{figure}%[tbhp]
\centering
\includegraphics[width=1\linewidth]{	Figures_for_MS/Fig2.png}
\caption{\footnotesize{Biological communities and their relationship with environmental variables in the summer months of the Hood Canal. \textbf{A} Constrained Analysis of Principal Coordinates (CAP) of Bray-Curtis dissimilarities among biological communities, as constrained by pH, temperature and salinity (arrows). \textbf{B} Most-likely cluster as a function of temperature and pH given a multinomial logistic model (salinity was uninformative, being tightly correlated with temperature). \textbf{C} Relative abundance (eDNA index; see \cite{kelly2019understanding}) of the taxa best distinguishing the three communities illustrated (SIMPER analysis). See Supporting Information for full analysis.}}
\label{fig:Fig2}
\end{figure}

@@

Diatoms (Bacillariophyta) show the steepest richness decline under future conditions (Figure 3 B.2); the probability of occurrence decreases markedly for diverse diatom taxa including both pennate (e.g., \textit{Nitzschia}) and centric (e.g., \textit{Coscinodiscus}, \textit{Stephanodiscus}) body forms. These declines in diatom richness were more accentuated at lower pH values and higher temperatures. Other taxa likely to find less-suitable habitat in the future include dinoflagellates \textit{Karlodinium} and Gymnodiniaceae sp.

@@

Given such heterogeneity in projections, gains and losses tend to balance one another out when looking at overall richness variation; we find no change in median richness for the year 2095 relative to the present in the Hood Canal (overall taxon richness by year, 95\% confidence interval in median species richness  -0.08, +0.04; Wilcoxon p $=$ 0.5); while higher diversity is expected in the San Juan Island in 2095 (increase in median species richness of 1.84-2.2, p $< 10^{-16}$). Diatoms, in particular, show small but significant declines in richness in the Hood Canal (0.47-0.57 species, p $< 10^{-16}$), while the changes on the San Juan Island are negligible (median change 0-0.08 p $=$ 0.04).  Dinoflagellates see their richness increase in both regions with the future scenario (median change 0.14-0.18 Hood Canal; 0.21-0.25 San Juan Island; p $< 10^{-16}$ for each).

%%%%%
 \begin{figure*}
 \centering
 \includegraphics[width=1\linewidth]{Figures_for_MS/Fig3.png}
\caption{\footnotesize{Forecasted changes in plankton in the Salish Sea for Dinoflagellates (panel \textbf{A}), Diatoms (\textbf{B}) and Haptophytes (\textbf{C}). Each panel shows:  (\textbf{1}) probability densities for the occurrence of selected taxa (species- and genus-level) for 2017 (blue) and 2095 (red); data are mean probabilities over 100 model draws, and variance in probability is due to differences in underlying environmental conditions.  (\textbf{2}) Changes in relative species richness within each phylum across the simulated scenario. (\textbf{3}) Relative taxon richness (raster color, warmer colors are more taxon-rich) for each of these same higher-taxa, for plausible ranges of pH and Temperature. Envelopes of observed (2017, blue) and modeled (2095, red) annual conditions in the Salish Sea shown for reference. Hood Canal and San Juan Island plotted separately to illustrate environmental differences between them.}}

>>>>

%% All changes due to new run of data are shown with *** before them for ease of search
Despite geographic proximity and similar overall species composition (127 of the 222 planktonic taxa were found in both regions and accounted for 98\% of the sequences), the areas under study -- San Juan Island and the Hood Canal -- varied widely in pH, temperature, and other environmental parameters (Figure 1C), with a smooth gradient in conditions along the Hood Canal, and San Juan Island more closely resembling full marine conditions. Different points along the environmental continuum simultaneously showed differences equivalent to those predicted between present-day and future oceans \cite{bopp2013multiple}.

Metabarcoding analysis of eDNA samples generated more than ***50.8M sequences from 778 samples. These samples represented biological and technical replicates from 86 unique sampling events. After bioinformatics quality-control \st{(See Methods)} the dataset included $\sim${}***45M sequences, from ***4849 unique amplicon sequence variants (ASVs). Of these, ***1364 ASVs (22.6M reads) could be annotated to a taxonomic level of Family or lower. These $\sim$ 500 taxa from 43 phyla were split according to their natural history and habitat (benthic vs. planktonic; see Supporting Information). Because we expect planktonic taxa to vary with water mass \cite{kelly2018tides} and therefore with bottle-sampled carbonate chemistry, here we focus on only the planktonic taxa (N = 221). These taxa showed a seasonal richness gradient between study areas, consistent with documented biodiversity clines in the area (Fig. 1B.1) \cite{dethier_decadal_2008}.

Bray-Curtis dissimilarities among samples revealed large differences in metabarcoding communities due to geographic Area ***(Hood Canal vs. San Juan Island; F-model = 1.6184; p $<$ 0.01). We therefore performed a constrained analysis of principal components (CAP) for samples within each Area, showing the differences among communities as a function of temperature, pH, and salinity (Fig. 2; for clarity, results for Hood Canal shown; full analysis in the Supporting Information).

@@

These communities were distinguished by changes in the relative abundances of a wide variety of taxa. In the Hood Canal for example, the community linked with colder water and higher-pH (community **2 in Fig. 2C) showed higher eDNA indexes of diatoms ***(\textit{Minutocellus polymorphus}), green algae  (\textit{Bathycoccus prasinos} and \textit{Micromonas pusilla}), and dinoflagellates like \textit{Karlodinium} sp than that on the low pH spectrum. In that community (cluster 1) \textit{Emiliania huxleyi}, \textit{Ditylum brightwellii} and the copepod \textit{Paracalanus} sp. C are more prevalent.  The community thriving in the high temperature range (community 3 in Fig. 2C) shows high values of \textit{Phaeocystis globosa},\textit{Ditylum brightwellii} among other species. Community 3 (Fig. 2A,B) and similar planktonic communities occupy the spectrum of environmental conditions most likely to be encountered in near-future climate scenarios as temperature rises and pH falls. For example, we expect the Hood Canal in 2095 to have the conditions in which community 3 is the most likely community 66\% of the time, an increase of 11\% compared to 2017(see Supplemental Material). On the other hand, conditions that make cluster 2 the most likely community drop from 9\% to 1\% of the time between 2017 and 2095.

@@

\begin{figure}%[tbhp]
\centering
\includegraphics[width=1\linewidth]{	Figures_for_MS/Fig2}
\caption{\footnotesize{Biological communities and their relationship with environmental variables in the summer months of the Hood Canal. \textbf{A} Constrained Analysis of Principal Coordinates (CAP) of Bray-Curtis dissimilarities among biological communities, as constrained by pH, temperature and salinity (arrows). \textbf{B} Most-likely cluster as a function of temperature and pH given a multinomial logistic model (salinity was uninformative, being tightly correlated with temperature). \textbf{C} Relative abundance (eDNA index; see \cite{kelly2019understanding}) of the taxa best distinguishing the three communities illustrated (SIMPER analysis). See Supporting Information for full analysis.}}
\label{fig:Fig2}
\end{figure}

@@

Diatoms (Bacillariophyta) show the steepest richness decline under future conditions (Figure 3 B.2); the probability of occurrence decreases markedly for diverse diatom taxa including both pennate (e.g., \textit{Nitzschia}) and centric (e.g., \textit{Coscinodiscus}, \textit{Stephanodiscus}) body forms. These declines in diatom richness were more accentuated at lower pH values and higher temperatures. Other taxa likely to find less-suitable habitat in the future include dinoflagellates \textit{Karlodinium} sp. and Gymnodiniaceae sp.

@@

Given such heterogeneity in projections, gains and losses tend to balance one another out when looking at overall richness variation; we find no change in median richness for the year 2095 relative to the present in the Hood Canal (overall taxon richness by year, 95\% confidence interval in median species richness  -0.08, +0.04; Wilcoxon p $=$ 0.5); while higher diversity is expected in the San Juan Island in 2095 (increase in median species richness of 1.84-2.2, p $< 10^{-16}$). Diatoms, in particular, show small but significant declines in richness in the Hood Canal (0.46-0.55 species, p $< 10^{-16}$), while the changes on the San Juan Island are negligible (median change 0-0.08 p $=$ 0.04).  Dinoflagellates see their richness increase in both regions with the future scenario (median change 0.14-0.18 Hood Canal; 0.21-0.25 San Juan Island; p $< 10^{-16}$ for each).

%%%%%
 \begin{figure*}
 \centering
 \includegraphics[width=1\linewidth]{Figures_for_MS/Fig3}
\caption{\footnotesize{Forecasted changes in plankton in the Salish Sea for Dinoflagellates (panel \textbf{A}), Diatoms (\textbf{B}) and Haptophytes (\textbf{C}). Each panel shows:  (\textbf{1}) probability densities for the occurrence of selected taxa (species- and genus-level) for 2017 (blue) and 2095 (red); data are mean probabilities over 100 model draws, and variance in probability is due to differences in underlying environmental conditions.  (\textbf{2}) Changes in relative species richness within each phylum across the simulated scenario. (\textbf{3}) Relative taxon richness (raster color, warmer colors are more taxon-rich) for each of these same higher-taxa, for plausible ranges of pH and Temperature. Envelopes of observed (2017, blue) and modeled (2095, red) annual conditions in the Salish Sea shown for reference. Hood Canal and San Juan Island plotted separately to illustrate environmental differences between them.}}

### 2020-09-24 Scripts/From.asv.to.data.plankton.Rmd
Rmarkdown shows consistency and figures are good to go. Latex document updated

```{r}
library (tidyverse)
library (here)
library (lubridate)
source(here("Scripts", "eDNAindex.R"))
ASV.table  <- read_csv(here("Input", "ASV_table_all_together.csv"))
Annotation <- read_csv(here("Input", "hash.annotated.csv"))
env.data   <- read_csv(here("Input", "env.data.updated.csv"))
```

@@

@@

env.data %>%  
   select( -contains("DO")) %>% 
   drop_na() %>% 
   right_join (ASV.plankton.event) %>% 
   unite (Site, Date, col = "event", sep = "_") %>% 
  write_csv(here("Input", "Combined_Biol_Env_Plankton.csv"))
```

>>>>

```{r}
library (tidyverse)
library (here)
library (lubridate)
source(here("Scripts", "eDNAindex.R"))
ASV.table  <- read_csv(here("Output", "ASV_table_all_together.csv"))
Annotation <- read_csv(here("Input", "hash.annotated.csv"))
env.data   <- read_csv(here("Input", "env.data.updated.csv"))
```

@@

# How many Hashes have family or more, how many reads
left_join(ASV.table, Annotation) %>% 
  unite(family, genus, species, col = "taxa", sep = "|") %>% 
  filter(taxa != "NA|NA|NA") %>% 
  summarise(n_distinct(Hash), sum(nReads), n_distinct(taxa))

@@

ASV.plankton.event %>%
  ungroup() %>% 
  summarise(n_distinct(taxa))

env.data %>%  
   select( -contains("DO")) %>% 
   drop_na() %>% 
   right_join (ASV.plankton.event) %>% 
   unite (Site, Date, col = "event", sep = "_") %>% 
  write_csv(here("Input", "Combined_Biol_Env_Plankton.csv"))

```

### 2020-09-24 Scripts/Manuscript.Main.Analysis.Rmd
Rmarkdown shows consistency and figures are good to go. Latex document updated

                          function(.x,.y) {
                            .y %>% select (Temperature, Salinity = Salinity.new, pH = pH_new) 
                            caps <-  capscale(formula = .x ~ Temperature + Salinity + pH ,
                                                 data = .y %>%
                                                   select(DIC = mean.DIC.new.sal, Salinity = Salinity.new, Temperature, pH = pH_new) %>% 
                                                   select_if(is.numeric) %>%
                                                   decostand(method = "standardize", na.action = na.omit),
                                                 distance = "jaccard", 
                                                binary = T)

@@

         ) -> Clusters.analysis.and.cap

# Write out clusers and CAP points, and raw env data

@@

```{r}
taxa.to.follow %>% 
  filter (average > 0.01) %>%  
  arrange(desc(average)) %>% 
  group_by(Area, Comparison, ava > avb) %>% 
    slice(1:5) %>%

  transmute(Area, Comparison, taxa = value, diff = avb - ava, average = case_when(ava > avb ~ average, TRUE ~ -average), sd) %>% 
  left_join(phylum.info %>% select(phylum,taxa)) %>% 
  separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
  mutate(label = case_when((species == "NA" & genus == "NA") ~ paste0(family, " sp."),
                           species == "NA" ~  paste0(genus, " sp."),
                           TRUE            ~ species)) %>%
  mutate(label2 = paste(phylum, label, sep = " | ")) %>% 
  mutate(Comparison = fct_relabel(Comparison, ~ str_replace(.,"_", " vs "))) %>% 
  nest(-Area) %>% 
  mutate(plot = map(data, function(.x){
    lims = c(-max(abs(.x$average) + (.x$sd)), max(abs(.x$average) + (.x$sd)))
    ggplot(.x, aes(y = average,
             x = fct_reorder(label2, phylum))) +
  geom_col(aes(fill = diff>0), size = 0.5, color = "black") +
      geom_errorbar(aes(ymin = average -sd, ymax = average + sd ), width = 0.25) +
  geom_hline(yintercept = 0, color = "black")+
  labs (x = "taxa",
        y = "Contribution to dissimilarity")+
  coord_flip(ylim=lims) +
    scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) +
  facet_wrap(~Comparison) +
  guides(fill = "none") +

@@

richness.by.phylum.by.point %>%
  filter(phylum == "Dinophyceae") %>%
  group_by(Year, Area) %>%
  summarise(median(richness))
explore.phylum("Dinophyceae")

#For example, we expect the Hood Canal to be T $>$ 12\degree C and pH $<$ {7.9} -- where community 3 is the most likely community -- more than four times as often in 2095 than in 2017 (4.9 \% vs 20.7\%) )

stds_to_value$toValue[[3]]
stds_to_value$toValue[[4]]

richness.by.phylum.by.point %>% 
  ungroup() %>% 
  dplyr::select(Area, Temperature, pH, Year) %>% 
  distinct() %>% 
  filter(Area == "Hood Canal") %>% 
  mutate(Temperature = 14.995 + 4.299*Temperature,
         pH = 8.2157 + 0.1813*pH) %>% 
  group_by(Year) %>% 
  add_count() %>% 
  summarise(condition = sum(Temperature > 12 & pH < 7.9),
            totalcount = n()) %>% 
  mutate(percent_time = 100 *condition / totalcount)
```

In general, the prevalence of cluster 3 in the Hood Canal will increase from 54.2% of the time to 75% of the time,
```{r echo = F}
# Add the cluster information to the projected scenario

@@

```{r echo = F}
#we find no change in median richness for the year 2095 relative to the present in the Hood Canal (overall taxon richness by year, averaged over all environmental conditions, Wilcoxon p $=$ 0.5);

wilcox.test(x = richness.by.point %>%
                  filter(Year == 2017, Area == "Hood Canal") %>% 
                    pull(richness),
            y = richness.by.point %>%
                  filter(Year == 2095, Area == "Hood Canal") %>% 
                    pull(richness), conf.int = T)

#while higher diversity is expected in the San Juan Island in 2095 (increase in median species richness of 2.1, Wilcoxon p $<$ 10 ^{-16}).

wilcox.test(x = richness.by.point %>%
                  filter(Year == 2017, Area == "San Juan Island") %>% 
                    pull(richness),
            y = richness.by.point %>%
                  filter(Year == 2095, Area == "San Juan Island") %>% 
                    pull(richness), conf.int = T)

#Diatoms, however, show small but significant declines in richness in the Hood Canal (0.5 species, Wilcoxon p $< 10^{-16}$),

wilcox.test(x = richness.by.phylum.by.point %>%
                  filter(Year == 2017, Area == "Hood Canal", phylum == "Bacillariophyta") %>% 
                    pull(richness),
            y = richness.by.phylum.by.point %>%
                  filter(Year == 2095, Area == "Hood Canal", phylum == "Bacillariophyta") %>% 
                    pull(richness), conf.int = T)

#while the changes on the San Juan Island are negligible (difference in median switch of 0.05, with  p $=$ 0.01).

wilcox.test(x = richness.by.phylum.by.point %>%
                  filter(Year == 2017, Area == "San Juan Island", phylum == "Bacillariophyta") %>% 
                    pull(richness),
            y = richness.by.phylum.by.point %>%
                  filter(Year == 2095, Area == "San Juan Island", phylum == "Bacillariophyta") %>% 
                    pull(richness), conf.int = T)

#Dinoflagellates see their richness increase in both regions with the future scenario(Wilcoxon p $< 10^{-16}$)

@@

wilcox.test(x = richness.by.phylum.by.point %>%
                  filter(Year == 2017, Area == "San Juan Island", phylum == "Dinophyceae") %>% 
                    pull(richness),
            y = richness.by.phylum.by.point %>%
                  filter(Year == 2095, Area == "San Juan Island", phylum == "Dinophyceae") %>% 
                    pull(richness),
            conf.int= TRUE)

#For example, in in the Hood Canal at present, we expect surface waters to have pH $<$ 7.9 and T $>$ 19$\degree$C only 1\% of the time. In 2095, we expect these conditions 6 times more frequently (i.e., 6\% of the time).

stds_to_value$toValue[[3]]
stds_to_value$toValue[[4]]

richness.by.phylum.by.point %>% 
  ungroup() %>% 
  dplyr::select(Area, Temperature, pH, Year) %>% 
  distinct() %>% 
  filter(Area == "Hood Canal") %>% 
  mutate(Temperature = 14.995 + 4.299*Temperature,
         pH = 8.2157 + 0.1813*pH) %>% 
  group_by(Year) %>% 
  add_count() %>% 
  summarise(condition = sum(Temperature > 19 & pH < 7.9),
            totalcount = n()) %>% 
  mutate(percent_time = condition / totalcount)

#At these values of T and pH, our model predicts the harmful \textit{Alexandrium} sp. to occur more often than not (mean frequency of occurrence $=$ 0.833)

expanded.future %>% 
  filter(Area == "Hood Canal", Year == 2095, TemperatureCelsius > 19, pH_units < 7.90, taxa == "Gonyaulacaceae|Alexandrium|NA") %>% 
  dplyr::select(present) %>% 
  summarise(mean(present))

#By contrast, the large centric diatom \textit{Coscinodiscus} -- a potentially key source of carbon for zooplankton and small fishes \cite{pasquaud2010determination, zamon2002tidal} with effects on dissolved oxygen and other water-column characteristics \cite{manabe1991bloom} -- occurs only one-third of the time under these same conditions (mean frequency $=$ 0.35)
expanded.future %>% 
  filter(Area == "Hood Canal", Year == 2095, TemperatureCelsius > 19, pH_units < 7.90, taxa == "Coscinodiscaceae|Coscinodiscus|Coscinodiscus wailesii") %>% 
  dplyr::select(present) %>% 
  summarise(mean(present))

# Responses in diatoms are not all in the same direction - Pseudonitzschia spp shows higher probability of presence at low pH values, and low pH values are more common in the 2095 scenario

expanded.future %>% 
  group_by(Area) %>% 
  filter(  str_detect(taxa, "Pseudo-nitzschia") ) %>% 
  ggplot(aes(x = pH_units, y = present)) +
  geom_smooth(aes(color = as.factor(Year), group = Year)) +
  facet_grid(~Area) + 
  ggtitle("Pseudo-nitzschia spp. response to pH")

  mutate (Area = str_replace(Area, replacement = "San Juan Island", pattern =  "SJI")) %>%
  select(Area, Year, Temperature = TemperatureCelsius,
         pH_new = pH_units) %>% 
  distinct() %>% 
    group_by(Area) %>% 
    nest() %>% 
    left_join(clusters) %>% 
    mutate(whichComm.res = map2(data,multi.mod,~ .x %>% 
                                  mutate(comm = as.numeric (apply(predict(.y, type = "probs", newdata = .x),1, which.max)) )
              ) ) %>% unnest(whichComm.res)
```

>>>>

                          function(.x,.y) {
                            .y %>% select (Temperature, Salinity = Salinity.new, pH = pH_new) 
                            caps <-  capscale(formula = .x ~ Temperature + Salinity + pH ,
                                                 data = .y %>%
                                                   select(DIC = mean.DIC.new.sal, Salinity = Salinity.new, Temperature, pH = pH_new) %>% 
                                                   select_if(is.numeric) %>%
                                                   decostand(method = "standardize", na.action = na.omit),
                                                 distance = "bray", 
                                                binary = F)

@@

         ) -> Clusters.analysis.and.cap
write_rds(Clusters.analysis.and.cap, here("Output", "CAP.halfway.rds"))
# Write out clusers and CAP points, and raw env data

@@

```{r}
taxa.to.follow %>% 
 # filter (average > 0.01) %>%  
  arrange(desc(average)) %>% 
  group_by(Area, Comparison, ava > avb) %>% 
    slice(1:5) %>%

  transmute(Area, Comparison, taxa = value, diff = avb - ava, average = case_when(ava > avb ~ average, TRUE ~ -average), sd) %>% 
  left_join(phylum.info %>% select(phylum,taxa)) %>% 
  separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
  mutate(label = case_when((species == "NA" & genus == "NA") ~ paste0(family, " sp."),
                           species == "NA" ~  paste0(genus, " sp."),
                           TRUE            ~ species)) %>%
  mutate(label2 = paste(phylum, label, sep = " | ")) %>% 
  mutate(Comparison = fct_relabel(Comparison, ~ str_replace(.,"_", " vs "))) %>% 
  nest(-Area) %>% 
  mutate(plot = map(data, function(.x){
    lims = c(-max(abs(.x$average) + (.x$sd)), max(abs(.x$average) + (.x$sd)))
    ggplot(.x, aes(y = average,
             x = fct_reorder(label2, label2))) +
  geom_col(aes(fill = diff>0), size = 0.5, color = "black") +
      geom_errorbar(aes(ymin = average -sd, ymax = average + sd ), width = 0.25) +
  geom_hline(yintercept = 0, color = "black")+
  labs (x = "taxa",
        y = "Contribution to dissimilarity")+
  coord_flip(ylim=lims) +
    scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) +
  facet_wrap(~Comparison) +
  guides(fill = "none") +

@@

richness.by.phylum.by.point %>%
  filter(phylum == "Dinophyceae") %>%
  group_by(Year, Area) %>%
  summarise(median(richness))
explore.phylum("Dinophyceae")
```

### Bonus track 3: Prevalence of clusters in future scenarios

For example, we expect the Hood Canal to be T $>$ 12\degree C and pH $<$ {7.9} -- where community 3 is the most likely community -- more than four times as often in 2095 than in 2017 (4.9 \% vs 20.7\%) )s
```{r}
input.for.left %>% 
  ungroup() %>% 
  select(Temperature = TemperatureCelsius,
         pH_new = pH_units, 
         Area,
         Year, sim.row) %>%
  distinct() %>% 
    group_by(Area) %>% 
    nest() %>% 
    left_join(clusters) %>% 
    mutate(whichComm.res = map2(data,multi.mod,~ .x %>% 
                                  mutate(comm = as.numeric (apply(predict(.y, type = "probs", newdata = .x),1, which.max)) ))) %>% 
  select(Area, whichComm.res) %>% 
  unnest(whichComm.res) %>% 
  add_count(Area, Year, name = "Tot") %>% 
  add_count(Area, Year, comm, name = "case.counts") %>% 
  group_by (Area, Year, comm) %>%

  summarise(percent_time = 100 * (case.counts/ Tot)) %>% distinct()
```

<!-- ```{r} -->
<!-- stds_to_value$toValue[[3]] -->
<!-- stds_to_value$toValue[[4]] -->

<!-- richness.by.phylum.by.point %>%  -->
<!--   ungroup() %>%  -->
<!--   dplyr::select(Area, Temperature, pH, Year) %>%  -->
<!--   distinct() %>%  -->
<!--   filter(Area == "Hood Canal") %>%  -->
<!--   mutate(Temperature = 14.995 + 4.299*Temperature, -->
<!--          pH = 8.2157 + 0.1813*pH) %>%  -->
<!--   group_by(Year) %>%  -->
<!--   add_count() %>%  -->
<!--   summarise(condition = sum(Temperature > 12 & pH < 7.9), -->
<!--             totalcount = n()) %>%  -->
<!--   mutate(percent_time = 100 *condition / totalcount) -->
<!-- ``` -->

In general, the prevalence of cluster 3 in the Hood Canal will increase from 55.6% of the time to 66.4% of the time,
```{r echo = F}
# Add the cluster information to the projected scenario

@@

**Overall richness, Hood Canal**
We find no change in median richness for the year 2095 relative to the present in the Hood Canal (overall taxon richness by year, averaged over all environmental conditions, Wilcoxon p $=$ 0.5)
```{r echo = F}
#;

wilcox.test(x = richness.by.point %>%
                  filter(Year == 2017, Area == "Hood Canal") %>% 
                    pull(richness),
            y = richness.by.point %>%
                  filter(Year == 2095, Area == "Hood Canal") %>% 
                    pull(richness), conf.int = T)
```

**Overall richness, San Juan Islands**
While higher diversity is expected in the San Juan Island in 2095 (increase in median species richness of 2.1, Wilcoxon p $<$ 10 ^{-16}).
```{r echo = F}
#

wilcox.test(x = richness.by.point %>%
                  filter(Year == 2017, Area == "San Juan Island") %>% 
                    pull(richness),
            y = richness.by.point %>%
                  filter(Year == 2095, Area == "San Juan Island") %>% 
                    pull(richness), conf.int = T)
```

**Diatom richness, Hood Canal**
Diatoms, however, show small but significant declines in richness in the Hood Canal (0.5 species, Wilcoxon p $< 10^{-16}$)
```{r echo = F}
wilcox.test(x = richness.by.phylum.by.point %>%
                  filter(Year == 2017, Area == "Hood Canal", phylum == "Bacillariophyta") %>% 
                    pull(richness),
            y = richness.by.phylum.by.point %>%
                  filter(Year == 2095, Area == "Hood Canal", phylum == "Bacillariophyta") %>% 
                    pull(richness), conf.int = T)
```

**Diatom richness, San Juan Islands**
While the changes on the San Juan Island are negligible (difference in median switch of 0.05, with  p $=$ 0.01).
```{r echo = F}
#

wilcox.test(x = richness.by.phylum.by.point %>%
                  filter(Year == 2017, Area == "San Juan Island", phylum == "Bacillariophyta") %>% 
                    pull(richness),
            y = richness.by.phylum.by.point %>%
                  filter(Year == 2095, Area == "San Juan Island", phylum == "Bacillariophyta") %>% 
                    pull(richness), conf.int = T)
```

**Dinoflagellates richness, Hood Canal**
Dinoflagellates see their richness increase in both regions with the future scenario(Wilcoxon p $< 10^{-16}$)
```{r echo = F}
#

@@

wilcox.test(x = richness.by.phylum.by.point %>%
                  filter(Year == 2017, Area == "San Juan Island", phylum == "Dinophyceae") %>% 
                    pull(richness),
            y = richness.by.phylum.by.point %>%
                  filter(Year == 2095, Area == "San Juan Island", phylum == "Dinophyceae") %>% 
                    pull(richness),
            conf.int= TRUE)
```

For example, in in the Hood Canal at present, we expect surface waters to have pH $<$ 7.9 and T $>$ 19$\degree$C less than 1\% of the time. In 2095, we expect these conditions 6 times more frequently (i.e., 6\% of the time).
```{r echo = F}

richness.by.phylum.by.point %>% 
  ungroup() %>% 
  dplyr::select(Area, Temperature, pH, Year) %>% 
  distinct() %>% 
  filter(Area == "Hood Canal") %>% 
  mutate(Temperature = 14.995 + 4.299*Temperature,
         pH = 8.2157 + 0.1813*pH) %>% 
  group_by(Year) %>% 
  add_count() %>% 
  summarise(condition = sum(Temperature > 19 & pH < 7.9),
            totalcount = n()) %>% 
  mutate(percent_time = 100*(condition / totalcount))
```

#### Species prevalence in future climate conditions
At these values of T and pH, our model predicts the harmful \textit{Alexandrium} sp. to occur more often than not (mean frequency of occurrence $=$ 0.833)
```{r echo = F}
#

expanded.future %>% 
  filter(Area == "Hood Canal", Year == 2095, TemperatureCelsius > 19, pH_units < 7.90, taxa == "Gonyaulacaceae|Alexandrium|NA") %>% 
  dplyr::select(present) %>% 
  summarise(mean(present))
```

By contrast, the large centric diatom \textit{Coscinodiscus} -- a potentially key source of carbon for zooplankton and small fishes \cite{pasquaud2010determination, zamon2002tidal} with effects on dissolved oxygen and other water-column characteristics \cite{manabe1991bloom} -- occurs only one-third of the time under these same conditions (mean frequency $=$ 0.35)
```{r echo = F}
#
expanded.future %>% 
  filter(Area == "Hood Canal", Year == 2095, TemperatureCelsius > 19, pH_units < 7.90, taxa == "Coscinodiscaceae|Coscinodiscus|Coscinodiscus wailesii") %>% 
  dplyr::select(present) %>% 
  summarise(mean(present))

# 
```

Responses in diatoms are not all in the same direction -  \textit{Pseudonitzschia} spp shows higher probability of presence at low pH values, and low pH values are more common in the 2095 scenario.

```{r echo = F}
expanded.future %>% 
  group_by(Area) %>% 
  filter(  str_detect(taxa, "Pseudo-nitzschia") ) %>% 
  ggplot(aes(x = pH_units, y = present)) +
  geom_smooth(aes(color = as.factor(Year), group = Year)) +
  facet_grid(~Area) + 
  ggtitle("Pseudo-nitzschia spp. response to pH")
